SunPy - Python for Solar Physics

### 2014-01-28 2-1-Map.tex
Some small typo fixed and latex stuff

\subsection{Map}
The purpose of the Map datatype is to represent 2D spatial data such as images of the Sun and 
inner heliosphere. It provides a wrapper around a data array and its associated
spatial coordinates and as well as meta data. The \texttt{Map} object provides 
convenience methods for many functions 
such as rotation and re-sampling as well as convenience visualisation 
functions, providing an easy to use as well as powerful interface. The \texttt{Map} object
also provides a convenient interface for loadings data from a variety of sources, typically
a FITS file as shown in Listing~\ref{code:aia_1}.

The architecture of the \texttt{Map} object consists of a template map called
\texttt{GenericMap} which is a subclass of \texttt{astropy.nddata.NDData}. \texttt{NDData} is a 
generic wrapper around a \texttt{ndarray} with a \texttt{.meta} attribute to store meta data.
As \texttt{NDData} is currently still in development, \texttt{GenericMap} does not make full
use of its capabilities but this inheritance structure provides 
for future integration of \texttt{astropy}. In order to provide instrument/
detector specific integration \texttt{GenericMap} is subclassed on creation. 
Each subclass of \texttt{GenericMap} can register with the \texttt{Map} creation function (i.e. factory)
by implementing a method that returns \texttt{True} if matching meta data 
for that instrument or detector is found in the file header. The \texttt{Map} factory 
will then automatically return an instance of the specific \texttt{GenericMap} 
subclass. As of this version SunPy has \texttt{Map} specialisations for the 
following instruments: 
\textit{YOHKOH/SXT},\textit{SOHO/EIT} and \textit{LASCO}, \textit{RHESSI}, 
\textit{STEREO/EUVI} and \textit{COR}, \textit{HINODE/XRT},
\textit{PROBA2/SWAP}, \textit{SDO/AIA} and \textit{HMI}, 
and \textit{IRIS} SJI frames.

The \texttt{Map} object stores standard meta data retrieved in the header of 
the image file, via the \texttt{.meta} attribute and provides convenience 
properties for more standard meta data, \textit{i.e.} \texttt{.instrument}, 
\texttt{.wavelength} or \texttt{.coordinate\_system}. 
Listing \ref{code:aia_1} also demonstrates the quick plot functionality of 
\texttt{Map}, see Section \ref{subsec:Viz} for more discussion of plotting.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> aiamap = sunpy.map.Map('aia_file.fits')
>>> aiamap.peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_map_example}
\end{center}
\caption{Demonstration of the \texttt{AIAMap} specialisation of 
\texttt{GenericMap}. The Map is created from an \textit{AIA} FITS file and the 
key meta data and array overview is printed. Then a quick view plot is created 
by using the \texttt{.peek()} method.}
\label{code:aia_1}
\end{listing}

As well as providing the base classes the map sub-package provides two 
collection classes, \texttt{CompositeMap} and \texttt{MapCube}, for 
temporally and spatially aligned data respectively.
\texttt{CompositeMap} provides methods for overlaying spatially aligned 
data, with support for visualisation of images and contour lines overlaid 
upon each other.
\texttt{MapCube} 
provides methods for animation of its series of \texttt{Map} objects. 
Listing~\ref{code:compmap_1} and \ref{code:mapcube_1} shows how to interact 
with these classes.
In addition, Listing~\ref{code:mapcube_2} demonstrates how the \texttt{MapCube} 
classes interacts with the \texttt{animation} module to save a video.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import matplotlib.pyplot as plt
>>> import sunpy.map
>>> compmap = sunpy.map.Map('aia_1600_image.fits', 'RHESSI_image.fits', 
...                         composite=True)
>>> compmap.set_levels(1, range(0,50,5), percent=True)
>>> compmap.set_colors(1, 'Reds_r')
#Plot the result and crop
>>> ax = plt.subplot()
>>> compmap.plot()
>>> ax.axis([200, 600, -600, -200])
>>> plt.show()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{comp_map_example}
\end{center}
\caption{Example demonstrating a CompositeMap plot, using contours and how 
SunPy integrates with matplotlib's pyplot functional interface.}
\label{code:compmap_1}
\end{listing}

%schriste - stuart, i agree if you want to remove the following two listings.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{python}
import sunpy.map

compmap = sunpy.map.Map('aia_lev1_171a_2014_01*fits', cube=True)
compmap.peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_cube_controls}
\end{center}
\caption{An example showing creation of a MapCube from a glob file search. The 
resultant plot makes use of matplotlib's interactive widgets to allow scrolling 
through the MapCube.}
\label{code:mapcube_1}
\end{listing}

\begin{listing}[h]
\begin{minted}[bgcolor=bg]{python}
import sunpy.map
from matplotlib import animation

mapc = sunpy.map.Map('aia_lev1_171a_2014_01*fits', cube=True)
anim = mapc.plot()
Writer = animation.writers['ffmpeg']
writer = Writer(fps=10, metadata=dict(artist='SunPy'), bitrate=1800)
anim.save('aia_cube.ogv')
\end{minted}
\caption{Example showing how to save a video animation from a MapCube, using 
matplotlib's animation framework.}
\label{code:mapcube_2}
\end{listing}

>>>>

\subsection{Map}\label{ssec:map}
The Map data type represents 2D spatial data such as images of the Sun and 
inner heliosphere. It provides a wrapper around a data array and its associated
spatial coordinates and other metadata. The \texttt{Map} class provides 
methods for typical operations on 2D data, 
such as rotation and re-sampling, and visualisation functionality.
The \texttt{Map} class
also provides a convenient interface for loading data from a variety of sources, typically
a FITS file as shown in Listing~\ref{code:aia_1}.

The architecture of the \texttt{Map} class consists of a template map called
\texttt{GenericMap}, which is a subclass of \texttt{astropy.nddata.NDData}. \texttt{NDData} is a 
generic wrapper around an \texttt{ndarray} with a \texttt{.meta} attribute to store metadata.
As \texttt{NDData} is currently still in development, \texttt{GenericMap} does not yet make full
use of its capabilities, but this inheritance structure provides 
for future integration of \texttt{Astropy}. In order to provide instrument- or
detector-specific integration, \texttt{GenericMap} is subclassed on creation. 
Each subclass of \texttt{GenericMap} can register with the \texttt{Map} creation function (\textit{i.e.}, factory)
by implementing a method that returns \texttt{True} if the metadata
in the file header matches
that instrument or detector.  This \texttt{Map} factory 
will then automatically return an instance of the specific \texttt{GenericMap} 
subclass. SunPy 0.4 has \texttt{Map} specialisations for the 
following instruments: 
\textit{YOHKOH/SXT}, \textit{SOHO/EIT} and \textit{LASCO}, \textit{RHESSI}, 
\textit{STEREO/EUVI} and \textit{COR}, \textit{HINODE/XRT},
\textit{PROBA2/SWAP}, \textit{SDO/AIA} and \textit{HMI}, 
and \textit{IRIS} SJI frames.

The \texttt{Map} class stores standard metadata retrieved from the header of 
the image file in the \texttt{.meta} attribute and provides convenience 
properties for commonly accessed metadata: \textit{e.g.}, \texttt{.instrument}, 
\texttt{.wavelength} or \texttt{.coordinate\_system}. 
Listing \ref{code:aia_1} demonstrates the quick-plot functionality of 
\texttt{Map}; see Section \ref{subsec:Viz} for more discussion of plotting.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> aiamap = sunpy.map.Map('aia_file.fits')
>>> aiamap.peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_map_example}
\end{center}
\caption{Example of the \texttt{AIAMap} specialisation of 
\texttt{GenericMap}. The Map is created from an \textit{AIA} FITS file,
and a quick-view plot is created.}
\label{code:aia_1}
\end{listing}

In addition to the data-type classes, the \texttt{map} sub-package provides two 
collection classes, \texttt{CompositeMap} and \texttt{MapCube}, for 
temporally and spatially aligned data respectively.
\texttt{CompositeMap} provides methods for overlaying spatially aligned 
data, with support for visualisation of images and contour lines overlaid 
upon each other.
\texttt{MapCube} 
provides methods for animation of its series of \texttt{Map} objects. 
Listings~\ref{code:compmap_1} and \ref{code:mapcube_1} show how to interact 
with these classes.
In addition, Listing~\ref{code:mapcube_2} shows how the \texttt{MapCube} 
classes interacts with the \texttt{animation} module to save a video.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import matplotlib.pyplot as plt
>>> import sunpy.map
>>> compmap = sunpy.map.Map('aia_1600_image.fits', 'RHESSI_image.fits', 
...                         composite=True)
>>> compmap.set_levels(1, range(0,50,5), percent=True)
>>> compmap.set_colors(1, 'Reds_r')
#Plot the result and crop
>>> ax = plt.subplot()
>>> compmap.plot()
>>> ax.axis([200, 600, -600, -200])
>>> plt.show()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{comp_map_example}
\end{center}
\caption{Example showing a \texttt{CompositeMap} plot and the integration with the \texttt{matplotlib.pyplot} interface.}
\label{code:compmap_1}
\end{listing}

%schriste - stuart, i agree if you want to remove the following two listings.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> compmap = sunpy.map.Map('aia_lev1_171a_2014_01*fits', cube=True)
>>> compmap.peek()
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{aia_cube_controls}
\end{center}
\caption{Example showing creation of a \texttt{MapCube} from a glob file search. The 
resultant plot makes use of \texttt{matplotlib}'s interactive widgets to allow scrolling 
through the \texttt{MapCube}.}
\label{code:mapcube_1}
\end{listing}

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> import sunpy.map
>>> from matplotlib import animation
>>> mapc = sunpy.map.Map('aia_lev1_171a_2014_01*fits', cube=True)
>>> anim = mapc.plot()
>>> Writer = animation.writers['ffmpeg']
>>> writer = Writer(fps=10, metadata=dict(artist='SunPy'), bitrate=1800)
>>> anim.save('aia_cube.ogv')
\end{minted}
% In the above example, is Writer, writer clear? 
% Could we add more metadata to the movie? obs_date, instrument, etc...?
\caption{Example showing how to save a video animation from a \texttt{MapCube}, using 
\texttt{matplotlib}'s animation framework.}
\label{code:mapcube_2}
\end{listing}

### 2014-01-28 2-2-Lightcurve.tex
removed that big git identifier tag thingy.
Some small typo fixed and latex stuff

Currently, the \texttt{LightCurve} class is compatible with the
following data sources: the GOES X-ray Sensor (XRS), PROBA2/LYRA, and
the SDO EUV Variability Experiment (EVE; only the level ``OCS'' and
average CSV files -- see \url{http://lasp.colorado.edu/home/eve/data/}
for more detail).  For each of these instruments, a subclass of the
\texttt{LightCurve} object is initialised
(\textit{e.g.}, \texttt{GOESLightCurve}) which inherits from
\texttt{LightCurve} but allows instrument-specific functionality to be
included.  Future developments will introduce support for additional
instruments and data products.  Since there is no established standard
as to how time-series data
should be stored and distributed, each SunPy \texttt{LightCurve}
object provides the ability to download its corresponding specific
data format in its constructor.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import lightcurve
>>> from sunpy.time import TimeRange
>>> goes = lightcurve.GOESLightCurve.create('2011-06-07 06:00',
...                                         '2011-06-07 10:00')
>>> goes.peek()
>>> trange = TimeRange('2011-06-07 06:00','2011-06-07 07:00')
%>>> goes_zoom = goes.truncate(trange)
% the above does not work!
>>> resampled_goes = goes.data.resample("1H")
\end{minted}
\begin{center}
\includegraphics[width=10cm]{goes_lightcurve.pdf}
\end{center}
\caption{Example retrieval of a GOES lightcurve for the time interval 06:00--08:00 UT on 
2011 June 7 using a time range, and the output of the \texttt{peek()} method.
A sub-interval is then extracted from the Lightcurve and finally the data is resampled
to 1 hour cadence.}
\label{code:goes_lc}
\end{listing}
%schriste - note to self, fix this example, add resampled points on top of plot?

>>>>

Currently, the \texttt{LightCurve} class is compatible with the
following data sources: the GOES X-ray Sensor (XRS), PROBA2/LYRA, and
the SDO EUV Variability Experiment (EVE; only the level ``OCS'' and
% Do we need to explain what OCS means?
average CSV files -- see \url{http://lasp.colorado.edu/home/eve/data/}
for more detail).  For each of these instruments, a subclass of the
\texttt{LightCurve} object is initialised
(\textit{e.g.}, \texttt{GOESLightCurve}) which inherits from
\texttt{LightCurve} but allows instrument-specific functionality to be
included.  Future developments will introduce support for additional
instruments and data products.  Since there is no established standard
as to how time-series data
should be stored and distributed, each SunPy \texttt{LightCurve}
object provides the ability to download its corresponding specific
data format in its constructor.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy import lightcurve
>>> from sunpy.time import TimeRange
>>> goes = lightcurve.GOESLightCurve.create('2011-06-07 06:00',
...                                         '2011-06-07 10:00')
>>> goes.peek()
>>> trange = TimeRange('2011-06-07 06:00','2011-06-07 07:00')
%>>> goes_zoom = goes.truncate(trange)
% the above does not work!
>>> resampled_goes = goes.data.resample("1H")
% Does 1h resample makes sense?? I vote for an example with some utility
\end{minted}
\begin{center}
\includegraphics[width=10cm]{goes_lightcurve.pdf}
\end{center}
\caption{Example retrieval of a GOES lightcurve for the time interval 06:00--08:00 UT on 
2011 June 7 using a time range, and the output of the \texttt{peek()} method.
A sub-interval is then extracted from the Lightcurve and finally the data is resampled
to 1 hour cadence.}
\label{code:goes_lc}
\end{listing}
%schriste - note to self, fix this example, add resampled points on top of plot?

### 2014-01-28 2-3-Spectra.tex
Some small typo fixed and latex stuff

\subsection{Spectra}\label{sec:spectra}
%schriste - this section needs major work

%ayshih - this whole paragraph can be deleted
Spectroscopy consists in the study of the radiative energy related to the wavelength.
An spectrum is normally obtained by observing a range of frequencies at the same time; 
it can be measured for example in radio wavelengths by a frequency sweep with 
a radio receiver or in visible and ultraviolet by the dispersion of the incident light 
through a diffraction grating or prism.
The analysis of the resultant spectrum can provide properties of the plasma observed 
such temperature, density, speed, etc.
Therefore, spectroscopy provides to the solar physicists invaluable information about 
the composition and the physical properties of the Sun.

SunPy aims to provide broad support for solar spectroscopy instruments, however the 
variety and complexity of such datasets makes it very challenging to add support to them all 
in a general way.
The \texttt{spectra} module implements a \texttt{Spectrum} class for 1D data
(intensity versus frequency) and a \texttt{Spectrogram} class for 2D data
(intensity versus time and frequency.
Each of these classes use a \texttt{numpy.ndarray} class as its \texttt{.data} attribute.
These two classes were implemented by funding provided by the Astrophysics Research 
Group at Trinity College Dublin.

As of SunPy 0.4, the \texttt{Spectrogram} class supports radio spectra from the e-Callisto 
solar radio spectrometer network (\url{http://www.e-callisto.org/})
and STEREO/SWAVES spectrograms.
As with other SunPy data types, the \texttt{Spectrogram} class has been
built so each instrument initialises using a subclass containing the instrument-specific 
functionalities.
The common functionality provided by the base \texttt{Spectrogram} class includes
reading data,
joining different time ranges and frequencies,
performing frequency-dependent background subtraction,
and convenient visualization and sampling of the data.

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram} object retrieves
from the online data archive the files that are between the time range specified and
the observatory of interest.  In this case the data is requested from \textit{BIR},
which is the code name of the
Rosse Observatory at Birr Castle in Ireland (\url{http://www.rosseobservatory.ie}).
When the data is requested using the \texttt{.from\_range} function, the object merges
into a single spectrogram all the files downloaded, each with 15 minutes observation data 
and in some cases with different files for different frequency ranges.  
In the example shown here BIR observed in two frequency ranges: 20--90\,MHz and 55--355\,MHz.
Since the data is not evenly spaced in the frequency range, the \texttt{Spectrogram} object
linearise the frequency axis for a better analisys, as shown in the resultant figures.
It is also shown in the example the effect of the implemented background substraction and
\texttt{.peek} method.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from SunPy.spectra.sources.callisto import CallistoSpectrogram
>>> tstart, tend = "2011-06-07T06:00:00", "2011-06-07T07:45:00"
>>> callisto = CallistoSpectrogram.from_range("BIR", tstart, tend)
>>> callisto_nobg = callisto.substract_bg()
>>> callisto_nobg.peek(vmin = 0)
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{callisto_nobg}
\end{center}
\caption{Example of how \texttt{CallistoSpectrogram} retrieves the data
for the time range and observatory requested, merges it all and removes the background
signal.}
\label{code:spectra}
\end{listing}

>>>>

\subsection{Spectra}\label{sec:spectra}
%schriste - this section needs major work
%ji - there is no example of a spectrum
%dps - I need to think in one...

SunPy aims to provide broad support for solar spectroscopy
instruments.  The variety and complexity of these instruments and
their resultant datasets makes it very challenging to support all in a
general way.  The \texttt{spectra} module implements a
\texttt{Spectrum} class for 1D data (intensity versus frequency) and a
\texttt{Spectrogram} class for 2D data (intensity versus time and
frequency).  Each of these classes use a \texttt{numpy.ndarray} class
as its \texttt{.data} attribute.  These two classes were implemented
by funding provided by the Astrophysics Research Group at Trinity
College Dublin, Ireland.

The \texttt{Spectrogram} class supports radio spectrograms from the e-Callisto 
solar radio spectrometer network (\url{http://www.e-callisto.org/})
and STEREO/SWAVES spectrograms.
As with other SunPy data types, the \texttt{Spectrogram} class has been
built so that each instrument initialises using a subclass containing the instrument-specific 
functionalities.
%The common functionality provided by the base \texttt{Spectrogram} class includes
%reading data,
%joining different time ranges and frequencies,
%performing frequency-dependent background subtraction,
%and convenient visualization and sampling of the data.

Listing \ref{code:spectra} shows how the \texttt{CallistoSpectrogram}
object retrieves spectrogram data in the time range specified taken at
the observatory of interest.  When the data is requested using the
\texttt{.from\_range} function, the object merges all the downloaded
files downloaded into a single spectrogram, across time and frequency.
In the example shown, data is observed in two frequency ranges, 
20--90\,MHz and 55--355\,MHz.  Since the data is not evenly spaced in
the frequency range, the \texttt{Spectrogram} object linearises the
frequency axis for a better analysis.  The example also demonstrates
the implemented background subtraction method.
%ji to DPS - what is the linearisation that is happening here?

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.spectra.sources.callisto import CallistoSpectrogram
>>> tstart, tend = "2011-06-07T06:00:00", "2011-06-07T07:45:00"
>>> callisto = CallistoSpectrogram.from_range("BIR", tstart, tend)
>>> callisto_nobg = callisto.subtract_bg()
>>> callisto_nobg.peek(vmin = 0)
\end{minted}
\begin{center}
\includegraphics[width=0.8\columnwidth]{callisto_nobg}
\end{center}
\caption{Example of how \texttt{CallistoSpectrogram} retrieves the
  data for the requested time range and observatory, merges it and
  removes the background signal.  The data requested -- 'BIR' -- is
  the code name of the Rosse Observatory at Birr Castle in Ireland
  (\url{http://www.rosseobservatory.ie})}
\label{code:spectra}
\end{listing}

### 2014-01-28 2-Data_Types.tex
changed the comment again
Some small typo fixed and latex stuff

\section{Core Data Types}
The primary focus of SunPy is to provide data structures 
that are specifically designed for the various kinds of solar physics data, namely
images, time series, and spectra. SunPy provides three core data
types using Python classes:
\texttt{Map} (2D spatial data),
\texttt{LightCurve} (1D temporal series)
and \texttt{Spectrum}/\texttt{Spectrogram} (1 and 2D spectra).

The classes provide access to the original data
along with associated metadata and provide appropriate convenience functions to
enable data analysis and visualization. For each of these classes, the data is
stored in the \texttt{.data} attribute while the metadata is stored 
in the \texttt{.meta} attribute. It is 
possible to instantiate the data types from various different 
sources: \textit{e.g.}, files, URLs, arrays or time ranges.  In order to provide
instrument-specific specialisation, all the core SunPy data types support specialisation;
\textit{e.g.}, \texttt{Map} has an \texttt{AIAMap} sub-type. 
Data visualisation is provided by two functions: \texttt{peek()}, for quick plotting,
and \texttt{plot()}, for more fine-grained control plotting that integrates with 
\texttt{matplotlib}.

@@

\subsection{Visualisation}
\label{subsec:Viz}
As has been demonstrated in this section, the core SunPy datatypes 
include visualisation code that is tailored to that data type. 
These visualisation methods all currently utilise the matplotlib 
package, and are designed in such a way that they integrate well with 
the \texttt{pyplot} functional interface of \texttt{matplotlib}, as demonstrated in 
Figure \ref{code:mpl_1}.

%schriste - I vote for removing all of the following code as being too down in the weeds
\begin{listing}[H]
\begin{minipage}{0.49\columnwidth}
\begin{minted}[bgcolor=bg]{python}
import sunpy.map
import matplotlib.pyplot as plt

>>>>

\section{Core Data Types}\label{sec:DataTypes}

The primary focus of SunPy is to provide data structures 
that are specifically designed for the primary varieties of solar physics data, namely
images, time series, and spectra. SunPy provides three core data
types using Python classes:
\texttt{Map} (2D spatial data),
\texttt{LightCurve} (1D temporal series)
and \texttt{Spectrum}/\texttt{Spectrogram} (1 and 2D spectra).

The classes provide access to the original data
along with associated metadata and provide appropriate convenience functions to
enable data analysis and visualisation. For each of these classes, the data is
stored in the \texttt{.data} attribute while the metadata is stored 
in the \texttt{.meta} attribute. It is 
possible to instantiate the data types from various different 
sources: \textit{e.g.}, files, URLs, arrays or time ranges.  In order to provide
instrument-specific specialisation, all the core SunPy data types support specialisation;
\textit{e.g.}, \texttt{Map} has an \texttt{AIAMap} sub-type. 
Data visualisation is provided by two functions: \texttt{peek()}, for quick plotting,
and \texttt{plot()}, for more fine-grained control plotting that integrates with 
\texttt{matplotlib}.
% LightCurve does NOT have a meta attribute.  Needs to be fixed in the
% re-write of LightCurve

@@

\subsection{Visualisation}
\label{subsec:Viz}
As has been demonstrated in this section, the core SunPy datatypes 
include visualisation code that is tailored to that data type. 
These visualisation methods all currently utilise the matplotlib 
package, and are designed in such a way that they integrate well with 
the \texttt{pyplot} functional interface of \texttt{matplotlib}, as demonstrated in 
Listing \ref{code:mpl_1}.

%schriste - I vote for removing all of the following code as being too
%down in the weeds
%ji -  I agree.  Do we even need to say much more than give a brief
%outline of the .peek and .plot methods?
\begin{listing}[H]
\begin{minipage}{0.49\columnwidth}
\begin{minted}[bgcolor=bg]{python}
import sunpy.map
import matplotlib.pyplot as plt

### 2014-01-28 3-1-VSO.tex
Some small typo fixed and latex stuff

The Virtual Solar Observatory (VSO) provides a single, standard query
interface to access solar data from different archives around the world
 \citep{hill2009}.
Data products can be requested for specific instruments or missions, but
can also be requested based on physical parameters of the data product such
as the wavelength range.
In addition to the VSO's primary, web-based interface
(\url{http://virtualsolar.org}), a SOAP (Simple Object Access Protocol)
service is also available.
SunPy's \texttt{vso} module accesses the VSO via the SOAP service using the
\texttt{suds} package.
This module was developed under the ESA's Summer of Code programme of 2011.

Listing~\ref{code:vso_query_simple} shows an example of how to query the VSO
using the \texttt{vso} module.
Each query is constructed by one or more attributes objects, where each
attribute object is a constraint on a parameter of the data set, such as the
time, instrument, wavelength, or  other properties.
The also shows how to download the data using the result
from the query. One can use the metadata associated with the data sets to specify how the
files will be named locally, such as creating subdirectories based on the
instrument names.

>>>>

The Virtual Solar Observatory (VSO) provides a single, standard query
interface to access solar data from different archives around the world
 \citep{hill2009}.
Data products can be requested for specific instruments or missions, but
can also be requested based on physical parameters of the data product such
as the wavelength range.
In addition to the VSO's primary, web-based interface
(\url{http://virtualsolar.org}), a SOAP (Simple Object Access Protocol)
service is also available.
SunPy's \texttt{vso} module accesses the VSO via the SOAP service using the
\texttt{suds} package.
This module was developed under the European Space Agency's Summer of Code in 
Space (ESA's SOCIS) of 2011.

Listing~\ref{code:vso_query_simple} shows an example of how to query the VSO
using the \texttt{vso} module.
Each query is constructed by one or more attributes objects, where each
attribute object is a constraint on a parameter of the data set, such as the
time, instrument, wavelength, or  other properties.
It also shows how to download the data using the result
from the query. One can use the metadata associated with the data sets to specify how the
files will be named locally, such as creating subdirectories based on the
instrument names.

### 2014-01-28 3-2-HEK.tex
Some small typo fixed and latex stuff

The Sun is an active star and exhibits a wide range of transient phenomena (e.g. flares, 
radio burst) on many 
different time-scales, length-scales and wavelengths. Observations and metadata 
concerning these phenomena are collected at the Heliophysics Event 
Knowledgebase (HEK).  Entries are generated both by automated algorithms and 
human observers.  Some of the information in the HEK reproduces feature/event 
data elsewhere (for example, the GOES flare catalog), and some is generated by 
the Solar Dynamics Observatory Feature Finding Team \cite{martens2012} to
populate the HEK.  The advantage of the HEK is it provides a homogeneous and 
well-described interface to a large 
amount of feature/event information of interest to the solar physics community.

SunPy accesses this information through the \texttt{hek} module, which was 
developed with support from the European Space Agency Summer of Code in 
Space (ESA-SOCIS) 2011.  The \texttt{hek} module makes use of the HEK API
(\url{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface?action=print}).
The current list of 
events maintained by the HEK and their properties can be found at 
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net import hek
>>> client = hek.HEKClient()
>>> tstart, tend = '2011/08/09 00:00:00', '2011/08/10 00:00:00'
>>> event_type = 'FL' # 'FL' indicates flare
>>> result = client.query(hek.attrs.Time(tstart,tend), 
...                       hek.attrs.EventType(event_type))
>>> len(result)
52
\end{minted}
\caption{Example usage of the \texttt{hek} module showing a simple HEK search for solar flares
which occurred on the 2011/08/09.}
\label{code:hek:simple}
\end{listing}

@@

HEK attributes differ from VSO attributes (Section \ref{ssec:vso}) in that many 
of them are wrappers that conveniently expose comparisons by overloading Python 
operators.  This allows filtering of the HEK entries by the properties of the 
event.  As was mentioned above, the HEK stores feature/event metadata obtained 
in different ways, known generally as {\it feature recognition methods} ( 
FRMs). Example \ref{code:hek:frm} filters the results of the previous result to 
return only those events that have the FRM 'SSW Latest Events'.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> result = client.query(hek.attrs.Time(tstart,tend), 
                      hek.attrs.EventType(event_type),
                      hek.attrs.FRM.Name == 'SSW Latest Events')
>>> len(result)
9
\end{minted}
\caption{An HEK query that returns only those flares that were
  detected by the 'SSW Latest Events' feature recognition method.}
\label{code:hek:frm}
\end{listing}
Multiple comparisons can be made by including more comma-separated
conditions on the attributes in the call to the HEK query method.

@@

The ability to use comparison and logical operators on HEK attributes allows 
the construction of queries of almost arbitrary complexity. 
It is also possible to use the 
results of a HEK query to download the corresponding data from the VSO by using 
SunPy's \texttt{H2VClient}.
%schriste - this is a weird way to end this section, either show how it is done or 
% do not mention it, i think.

>>>>

The Sun is an active star and exhibits a wide range of transient phenomena (e.g. flares, 
radio bursts) on many 
different time-scales, length-scales and wavelengths. Observations and metadata 
concerning these phenomena are collected at the Heliophysics Event 
Knowledgebase (HEK).  Entries are generated both by automated algorithms and 
human observers.  Some of the information in the HEK reproduces feature/event 
data elsewhere (for example, the GOES flare catalog), and some is generated by 
the Solar Dynamics Observatory Feature Finding Team \cite{martens2012} to
populate the HEK.  The advantage of the HEK is it provides an homogeneous and 
well-described interface to a large 
amount of feature/event information of interest to the solar physics community.

SunPy accesses this information through the \texttt{hek} module, which was 
developed with support from the 
ESA-SOCIS 2011.  The \texttt{hek} module makes use of the HEK API
(\url{http://vso.stanford.edu/hekwiki/ApplicationProgrammingInterface?action=print}).
The current list of 
events maintained by the HEK and their properties can be found at 
\url{http://www.lmsal.com/hek/VOEvent_Spec.html}.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net import hek
>>> client = hek.HEKClient()
>>> tstart, tend = '2011/08/09 00:00:00', '2011/08/10 00:00:00'
>>> event_type = 'FL' # 'FL' indicates flare
>>> result = client.query(hek.attrs.Time(tstart,tend), 
...                       hek.attrs.EventType(event_type))
>>> len(result)
52
\end{minted}
\caption{Example usage of the \texttt{hek} module showing a simple HEK search for solar flares
which occurred on August 9th, 2011.}
\label{code:hek:simple}
\end{listing}

@@

HEK attributes differ from VSO attributes (Section \ref{ssec:vso}) in that many 
of them are wrappers that conveniently expose comparisons by overloading Python 
operators.  This allows filtering of the HEK entries by the properties of the 
event.  As was mentioned above, the HEK stores feature/event metadata obtained 
in different ways, known generally as {\it feature recognition methods} ( 
FRMs). Example in Listing~\ref{code:hek:frm} filters the results of the previous result to 
return only those events that have the FRM 'SSW Latest Events'.
\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> result = client.query(hek.attrs.Time(tstart,tend), 
...                       hek.attrs.EventType(event_type),
...                       hek.attrs.FRM.Name == 'SSW Latest Events')
>>> len(result)
9
\end{minted}
\caption{An HEK query that returns only those flares that were
  detected by the 'SSW Latest Events' feature recognition method.}
\label{code:hek:frm}
\end{listing}
Multiple comparisons can be made by including more comma-separated
conditions on the attributes in the call to the HEK query method.

@@

The ability to use comparison and logical operators on HEK attributes allows 
the construction of queries of almost arbitrary complexity. 
It is also possible to use the 
results of a HEK query to download the corresponding data from the VSO by using 
SunPy's \texttt{H2VClient}.
%schriste - this is a weird way to end this section, either show how it is done or 
% do not mention it, i think.
%dps - I agree.
